Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2017-11-07 manuscript/usses_water_ms.tex
starts revisions; adds new appendix 4; fixes figures after bug in NDVI calcs

\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
regression model for observation \emph{i}, which is associated with plot
\emph{j} and treatment year \emph{t}. \(\boldsymbol{\beta}\) is the
vector of coefficients for the fixed effects in the design matrix
\(\textbf{X}\). Each row of the design matrix represents a single
observation (\(\textbf{x}_i\)) and is a vector with the following
elements: 1 for the intercept, a binary 0 or 1 if the treatment is
``drought'', a binary 0 or 1 if the treatment is ``irrigation'', the
scaled value of VWC, binary ``drought'' value times VWC, and binary
``irrigation'' value times VWC. Thus, our model treats ``control''
observations as the main treatment and then estimates intercept and
slope offsets for the ``drought'' and ``irrigation'' treatments. We use
our model to test two statistical hypotheses based on the questions in
our Introduction:

@@

We include two random effects to account for the fact that observations
within plots and years are not independent. Specifically, we include
plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
slope terms and year-specific intercept offsets (\(\eta_t\)). The
covariate vector \(\textbf{z}_i\) for each observation \emph{i} has two
elements: a 1 for the intercept and the scaled value of VWC for that
plot and year. The plot-specific coefficients are modeled
hierarchically, where plot level coefficients are drawn from a
multivariate normal distribution with mean 0 and a variance-covariance
structure that allows the intercept and slope terms to be correlated:

@@

\noindent{}where \(\Sigma\) is the variance-covariance matrix and
\emph{j(i)} reads as ``plot \emph{j} associated with observation
\emph{i}''. The random year effects (\(\boldsymbol{\eta}\)) are drawn
from a normal prior with mean 0 and standard deviation
\(\sigma_{\text{year}}\), which was drawn from a half-Cauchy
distribution. A full description of our model and associated R
\citep{R2016} code is in Appendix 2.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(Table 1; Fig. 3). The effect of soil moisture for each treatment is
associated with high uncertainty, however, with 95\% Bayesian credible
intervals that overlap zero (Table 1). Although the parameter estimates
for the effect of soil moisture overlap zero, the posterior
distributions of the slopes all shrank and shifted to more positive
values relative to the prior distributions (Fig. A2-2), which indicates
the data did influence parameter estimates beyond the information from
the uninformative priors. Ecosystem functional response was similar
among treatments (Table 1; Fig. 3B), but there is evidence that the
slope for the irrigation treatment is less than the slope for the
control treatment. This evidence comes from interpreting the posterior
distribution of the slope offset for the irrigation treatment, from
which we calculate a 99\% one-tailed probability that the estimate is
less than zero (Fig. 3A, right panel). There was no evidence that the
treatment effects became more important over time because there was no
directional trend in the random year effects (Fig. A2-3).

@@

First, it could be that our precipitation manipulations were not large
enough to induce a response. That is, a 50\% decrease in any given year
may not be abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our treatments \emph{should} have been large
enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site (Fig. A3-3). The 1\%
quantile of precipitation at our site is 110 mm, a 47\% reduction from
the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
growing season precipitation (Appendix 3). Thus, our drought treatment
represented extreme precipitation amounts, especially in years where
ambient precipitation was below average (Fig. 2A). The irrigation
treatment may not qualify as extreme, yet that is the treatment where we
did observe an effect (Fig. 3A).

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.22 & 0.16 & -1.25 & 2.16 \\ 
  Intercept & Drought & 0.64 & 0.44 & -1.37 & 3.79 \\ 
  Intercept & Irrigation & -0.51 & -0.37 & -2.99 & 1.26 \\ 
  Slope & Control & 1.31 & 1.14 & -0.30 & 3.83 \\ 
  Slope & Drought & 1.18 & 1.00 & -0.62 & 4.12 \\ 
  Slope & Irrigation & 0.83 & 0.68 & -0.61 & 3.12 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  21 &   2 & 1.02 & 0.44 \\ 
  2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.10 & 0.32 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.23 & 0.14 \\ 
  2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
  2014 & adonis &  22 &   2 & 0.95 & 0.58 \\ 
  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
  2015 & adonis &  21 &   2 & 1.05 & 0.40 \\ 
  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
  2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

>>>>

\noindent{}where \(\mu_{i}\) is the deterministic prediction from the
regression model for observation \(i\), which is associated with plot
\(j\) and treatment year \(t\). \(\boldsymbol{\beta}\) is the vector of
coefficients for the fixed effects in the design matrix \(\textbf{X}\).
Each row of the design matrix represents a single observation
(\(\textbf{x}_i\)) and is a vector with the following elements: 1 for
the intercept, a binary 0 or 1 if the treatment is ``drought'', a binary
0 or 1 if the treatment is ``irrigation'', the scaled value of VWC,
binary ``drought'' value times VWC, and binary ``irrigation'' value
times VWC. Thus, our model treats ``control'' observations as the main
treatment and then estimates intercept and slope offsets for the
``drought'' and ``irrigation'' treatments. We use our model to test two
statistical hypotheses based on the questions in our Introduction:

@@

We include two random effects to account for the fact that observations
within plots and years are not independent. Specifically, we include
plot-specific offsets (\(\boldsymbol{\gamma}\)) for the intercept and
slope terms and year-specific intercept offsets (\(\eta_t\)). The
covariate vector \(\textbf{z}_i\) for each observation \(i\) has two
elements: a 1 for the intercept and the scaled value of VWC for that
plot and year. The plot-specific coefficients are modeled
hierarchically, where plot level coefficients are drawn from a
multivariate normal distribution with mean 0 and a variance-covariance
structure that allows the intercept and slope terms to be correlated:

@@

\noindent{}where \(\Sigma\) is the variance-covariance matrix and
\(j(i)\) reads as ``plot \(j\) associated with observation \(i\)''. The
random year effects (\(\boldsymbol{\eta}\)) are drawn from a normal
prior with mean 0 and standard deviation \(\sigma_{\text{year}}\), which
was drawn from a half-Cauchy distribution. A full description of our
model and associated R \citep{R2016} code is in Appendix 2.

@@

Cumulative March-June soil moisture had a weak positive effect on ANPP
(Table 1; Fig. 3B). The effect of soil moisture for each treatment is
associated with high uncertainty, however, with 95\% Bayesian credible
intervals that overlap zero (Table 1). Although the parameter estimates
for the effect of soil moisture overlap zero, the posterior
distributions of the slopes all shrank and shifted to more positive
values relative to the prior distributions (Fig. A2-2), which indicates
the data did influence parameter estimates beyond the information from
the uninformative priors. Ecosystem functional response was similar
among treatments (Table 1; Fig. 3B), but there is evidence that the
slope for the irrigation treatment is less than the slope for the
control treatment. This evidence comes from interpreting the posterior
distribution of the slope offset for the irrigation treatment, from
which we calculate a \new{96}\% one-tailed probability that the estimate
is less than zero (Fig. 3A, right panel).

@@

First, it could be that our precipitation manipulations were not large
enough to induce a response. That is, a 50\% decrease in any given year
may not be abnormal given our site's historical range of variability
\citep{Knapp2017}. We cannot definitively rule out this possibility, but
we have reason to believe our treatments \emph{should} have been large
enough. Using the methods described by \citet{Lemoine2016}, we
calculated the percent reduction and increase of mean growing season
precipitation necessary to reach the 1\% and 99\% extremes of the
historical precipitation regime at our site (Fig. A4-1). The 1\%
quantile of precipitation at our site is 110 mm, a 47\% reduction from
the mean, and the 99\% quantile is 414 mm, a 77\% increase from mean
growing season precipitation (Appendix 4). Thus, our drought treatment
represented extreme precipitation amounts, especially in years where
ambient precipitation was below average (Fig. 2A). The irrigation
treatment may not qualify as extreme, yet that is the treatment where we
did observe an effect (Fig. 3A).

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.16 & 0.14 & -1.22 & 1.69 \\ 
  Intercept & Drought & 0.13 & 0.05 & -1.69 & 2.39 \\ 
  Intercept & Irrigation & -0.24 & -0.18 & -2.11 & 1.53 \\ 
  Slope & Control & 0.94 & 0.86 & -0.51 & 2.69 \\ 
  Slope & Drought & 0.65 & 0.57 & -0.99 & 2.62 \\ 
  Slope & Irrigation & 0.72 & 0.65 & -0.54 & 2.26 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  21 &   2 & 1.02 & 0.44 \\ 
  2011 & betadisper &  21 &   2 & 2.23 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.10 & 0.34 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.23 & 0.16 \\ 
  2013 & betadisper &  22 &   2 & 0.28 & 0.76 \\ 
  2014 & adonis &  22 &   2 & 0.95 & 0.56 \\ 
  2014 & betadisper &  22 &   2 & 0.35 & 0.71 \\ 
  2015 & adonis &  21 &   2 & 1.05 & 0.37 \\ 
  2015 & betadisper &  21 &   2 & 3.01 & 0.07 \\ 
  2016 & adonis &  21 &   2 & 1.07 & 0.36 \\ 
  2016 & betadisper &  21 &   2 & 0.50 & 0.62 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}
